Improving schedulability rate of EDF when taking into account the preemption cost in uniprocessor periodic systems

### 2013-10-08 python-simulation/paper-paedf/paper-paedf.tex
added more example to the article about examples

\documentclass[a4paper,10pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{charter}
\usepackage{listings}
\usepackage{color}
\usepackage{url}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{hyperref}
\usepackage{graphicx}
%\usepackage[section]{placeins}

@@

        We use an extension of Liu and Layland model presented in \cite{Liu:2000:RS:518501} to formalize real-time systems. This model represents such systems by a set of \textbf{tasks} generating \textbf{jobs}, where a job is a whole entity of computation with a given time of arrival and a given deadline.\\

@@

        Similarly, we differentiate between \textbf{synchronous system} (where $O_i = O_j \; \forall i,j$) and asynchronous systems (which are not synchronous).

@@

        The authors of \cite{liu1973scheduling} have shown that the EDF (Earliest Deadline First) scheduling policy, in which jobs are prioritized according to their absolute deadline, is optimal, meaning that any feasible system is schedulable by EDF.\\

@@

        \begin{figure}[h]
        \begin{center}
            \includegraphics[width=\textwidth]{figs/atomicpreemption_example.png}
            \caption{Schedule of a system with $\alpha=2$. At t $t=6$, the preemption recovery period is not interrupted even though a job with higher priority is activated}
            \label{fig:prp}
        \end{center}
        \end{figure}

        Meumeu is FTP-optimal (equivalent to ExhaustiveFTP)

\section{Performance of EDF with preemption cost}
% Rename section to something like "Old results do not old, bitch"

@@

        \begin{figure}[h]
        \begin{center}
            \includegraphics{figs/edfNonOptimal_EDF.png}
            \caption{A system scheduled by EDF where a deadline is missed}
            \label{fig:edfnonoptimal_edf}
        \end{center}
        \end{figure}

@@

        \begin{figure}[h]
        \begin{center}
            \includegraphics{figs/edfNonOptimal_PALLF.png}
            \caption{The system of Fig.~\ref{fig:edfnonoptimal_edf} with a different schedule in which no deadline is missed}
            \label{fig:edfnonoptimal_pallf}
        \end{center}
        \end{figure}

@@

        When scheduling this system with EDF, we see that at instant $O_{max} + 2 \cdot H$, the system has not reached a periodic behavior yet. This only happens at instant $O_{max} + 4 \cdot H$.\\

        Figure.

@@

        And consider the following schedule which shows that it is feasible with an idling scheduling policy.

        [figure]

        However, it is easy to see that no non-idling algorithm can successfully schedule this system: at $t=0$, a non-idling algorithm has to choose between the first job of $\tau_1$ and $\tau_2$ for execution, whichever is chosen will then be preempted by the first job of $\tau_3$ (if not, it will miss its deadline in $t=2$). Then, at $t=2$, there will be $3+2+2=7$ units of computation left to compute before the deadline at $t=8$, the system will thus certainly miss a deadline.

@@

        \subsubsection{u < 0.69 is not sufficient}

        \subsubsection{Require Claivoyance}

        \subsubsection{Preemption can happen anywhere}

>>>>

\documentclass[a4paper,10pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{charter}
\usepackage{listings}
\usepackage{color}
\usepackage{url}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{float}
%\usepackage[section]{placeins}

@@

        We use an extension of Liu and Layland model presented in \cite{Liu:2000:RS:518501} to formalize real-time systems. This model represents such systems by a set of \textbf{tasks} generating \textbf{jobs}, where a job is a whole entity of computation with a given time of arrival and a given deadline.

@@

        Similarly, we differentiate between \textbf{synchronous system} (where $O_i = O_j \; \forall i,j$) and asynchronous systems (which are not synchronous).\\

        To summarize, Fig.~\ref{fig:rt_ex} shows an example of real-time system in our model.

        \begin{figure}[H]
        \begin{center}
            \includegraphics{figs/RTsystem_example.png}
            \caption{A task system. Annotations will be removed in later figures}
            \label{fig:rt_ex}
        \end{center}
        \end{figure}

@@

        The authors of \cite{liu1973scheduling} have shown that the EDF (Earliest Deadline First) scheduling policy, in which jobs are prioritized according to their absolute deadline, is optimal, meaning that any feasible system is schedulable by EDF.

@@

        \begin{figure}[H]
        \begin{center}
            \includegraphics[width=\textwidth]{figs/atomicpreemption_example.png}
            \caption{Schedule of a system with $\alpha=2$. At t $t=6$, the preemption recovery period is not interrupted even though a job with higher priority is activated}
            \label{fig:prp}
        \end{center}
        \end{figure}

        The authors of \cite{meumeu2007extending} (TOCHECK) propose an unnamed (we use PA-RMA to refer to it) fixed-priority-optimal scheduling policy for synchronous implicit deadline systems.

\section{Performance of EDF with preemption cost}
% Rename section to something like "Old results do not hold, bitch"

@@

        \begin{figure}[H]
        \begin{center}
            \includegraphics{figs/edfNonOptimal_EDF.png}
            \caption{A system scheduled by EDF where a deadline is missed}
            \label{fig:edfnonoptimal_edf}
        \end{center}
        \end{figure}

@@

        \begin{figure}[H]
        \begin{center}
            \includegraphics{figs/edfNonOptimal_PALLF.png}
            \caption{The system of Fig.~\ref{fig:edfnonoptimal_edf} with a different schedule in which no deadline is missed}
            \label{fig:edfnonoptimal_pallf}
        \end{center}
        \end{figure}

@@

        When scheduling this system with EDF, we see that at instant $O_{max} + 2 \cdot H$, the system has not reached a periodic behavior yet. This only happens at instant $O_{max} + 4 \cdot H$, as seen on Fig.~\ref{fig:edf_longtransitive}.\\

        \begin{figure}[H]
        \begin{center}
            \centerline{\includegraphics[width=1.4\textwidth]{figs/longTransitive2_EDF.png}}
            \caption{Scheduling of a system where the periodic behavior does not happen until $O_{max} + 4 \cdot H$}
            \label{fig:edf_longtransitive}
        \end{center}
        \end{figure}

@@

        And consider the following schedule of fig.~\ref{fig:mustidle_pallf} which shows that it is feasible with an idling scheduling policy.

        \begin{figure}[H]
        \begin{center}
            \includegraphics[scale=0.7]{figs/mustIdle_PALLF.png}
            \caption{A system scheduled by an idling algorithm. $t=0$ is an idle instant even though a job of $\tau_1$ is active}
            \label{fig:mustidle_pallf}
        \end{center}
        \end{figure}

        However, it is easy to see that no non-idling algorithm can successfully schedule this system: at $t=0$, a non-idling algorithm has to choose between $j_{1,1}$ and $j_{2,1}$ for execution, whichever is chosen will then be preempted by $j_{3,1}$ (if not, it will miss its deadline in $t=2$).\\

        Which means that at $t=2$, there will either be \mbox{$(c_{1,1} - 1) +c_{2,1} + \alpha = 7$} or \mbox{$c_{1,1} + (c_{2,1} -1) + \alpha = 7$} units of computation left to compute (including preemptions recovery) before the deadline at $t=8$, the system will thus certainly miss a deadline.

@@

        \begin{center}
            \begin{tabular}{|r|c|c|c|c|c|}
                \hline
                            & $O_i$ & $C_i$ & $D_i$ & $T_i$ & $\alpha_i$ \\ \hline
                $\tau_1$    & 0     & 4     & 8    & 8     & 1     \\ \hline
                $\tau_2$    & 0     & 1     & 5    & 8     & 1     \\ \hline
                $\tau_3$    & 3     & 1     & 1    & 8     & 1     \\ \hline
                $\tau_4$    & 5     & 1     & 1    & 8     & 1     \\ \hline
            \end{tabular}
        \end{center}

        And a successful schedule of it in Fig.~\ref{fig:dponly_pallf}.

        \begin{figure}[H]
        \begin{center}
            \includegraphics[scale=0.7]{figs/dponly_palff.png}
            \caption{A system scheduled by a dynamic algorithm. The priorities of $j_{1,1}$ and $j_{2,1}$ change between $t=0$ and $t=4$}
            \label{fig:dponly_pallf}
        \end{center}
        \end{figure}

        Furthermore, it can be shown that the schedule of fig.~\ref{fig:dponly_pallf} is the only successful schedule possible. However no fixed-priority algorithm (either at the task of the job level) can produce it, which shows that dynamic algorithms are necessary when considering the preemption cost.

        % \subsubsection{$u < 0.69$ is not sufficient}

        \subsubsection{Require clairvoyance}

        Consider the following the following system

        \begin{center}
            \begin{tabular}{|r|c|c|c|c|c|}
                \hline
                            & $O_i$ & $C_i$ & $D_i$ & $T_i$ & $\alpha_i$ \\ \hline
                $\tau_1$    & 22    & 2     & 2    & 24    & 1     \\ \hline
                $\tau_2$    & 0     & 5     & 12   & 12    & 1     \\ \hline
                $\tau_3$    & 4     & 5     & 6    & 12    & 1     \\ \hline
                $\tau_4$    & 9     & 1     & 1    & 24    & 1     \\ \hline
            \end{tabular}
        \end{center}

        And a successful schedule of it

        \begin{figure}[H]
        \begin{center}
            \includegraphics[scale=0.7]{figs/clairvoyance_example.png}
            \caption{A system scheduled by a clairvoyant algorithm. Different decisions have to be taken at $t=4$ and $t=16$ even though the state of the system is identical}
            \label{fig:clairvoyance}
        \end{center}
        \end{figure}

        \subsubsection{Preemption at non-arrivals}

        As we have shown previously, an optimal algorithm has to update priority of the jobs dynamically. This can be very heavy if done at each instant, and one could wonder if updating the priority only at arrivals of job would be sufficient. We show by an example that this is not the case.\\

        \begin{center}
            \begin{tabular}{|r|c|c|c|c|c|}
                \hline
                            & $O_i$ & $C_i$ & $D_i$ & $T_i$ & $\alpha_i$ \\ \hline
                $\tau_1$    & 0     & 3     & 6    & 6     & 1     \\ \hline
                $\tau_2$    & 3     & 1     & 1    & 6     & 1     \\ \hline
                $\tau_3$    & 1     & 1     & 3    & 6     & 1     \\ \hline
            \end{tabular}
        \end{center}

        And the only valid schedule of it

        \begin{figure}[H]
        \begin{center}
            \includegraphics[scale=0.7]{figs/mpana.png}
            \caption{A valid schedule of a system. Note that the preemption occurs at $t=2$ which is not a job arrival time}
            \label{fig:mpana}
        \end{center}
        \end{figure}

        Indeed, if the preemption of $j_{1,1}$ happens at $t=1$, $j_{3,1}$ will finish its execution at $t=1$ and the instant $t=3$ will be used for a preemption recovery period which will be immediately followed by another preemption, leading to a deadline miss at $t=6$. Similarly, if $j_{1,1}$ continues its execution without preemption, either $j_{2,1}$ or $j_{3,1}$ will have to miss its deadline at $t=4$. The proposed schedule is thus the only valid and require a priority update at $t=2$.
